Conditions for $d$-local-connectedness of boundaries of hyperbolic groups

### 2020-02-05 paper.tex
Changed Delta^d+1 to Delta^2 in section 4

  We now bound the distance $\dist(y_n, y_{n+1})$. To start with, $\dist(y_n,
  i^n_{n+1} y_n) \leq 2\delta+1$. Let $R$ be $3\cdot 2^{k-1}$, so that any two
  vertices in the 1-skeleton of $\sd^k \Delta^{d+1}$ are joined by a path
  traversing at most $R$ edges. Let $\sigma$ be the minimal simplex of $K_n$
  containing $i^m_n(x)$. Then $y_n$ is a vertex of $\sigma$ and $y_{n+1}$ is a
  vertex of the image of $\sigma$ under $i^n_{n+1}$, so $\dist(i^n_{n+1} y_n,
  y_{n+1}) \leq RD$.  Putting this together,
  \begin{equation*}
      \dist(y_n, y_{n+1}) \leq 2\delta + 1 + RD
  \end{equation*}

@@

\begin{proposition}\label{prop:basic_homotopies}
  Let $C$ be a constant. Then there exists a number $m_0$ and a function $\mu
  \from \posreals\to\posreals$ such that $\mu(\eta)\to0$ as $\eta\to0$ with the following
  property. Let $n \geq m_0$ and let $f_0$ and $f_1$ be any maps from
  $\boundary\Delta^{d+1}$ to $K_n$ satisfying the following conditions:
  \begin{enumerate}
    \item the maps $f_0$ and $f_1$ are affine on some subdivision of
      $\boundary\Delta^{d+1}$;
    \item for any $x \in \boundary\Delta^{d+1}$, the distance 
      $\newdist(f_0(x), f_1(x))$ is at most $C$;
    \item each of $f_0$ and $f_1$ admits a factoring.
  \end{enumerate}
  Then there is a homotopy $h$ from $i^n_\infty\composed f_0$ to
  $i^n_\infty\composed f_1$ satisfying a uniform continuity condition: for any
  $x \in \boundary\Delta^{d+1}$ and any $s$ and $t$ in $[0,1]$,
  \begin{equation*}
    \rho(h(x,s), h(x,t)) \leq \mu(\abs{s-t})a^{-n}.
  \end{equation*}
\end{proposition}

@@

  Let $n \geq m_0$.  Let $J$ be a subdivision of $\boundary\Delta^{d+1}$ such
  that $f_0$ and $f_1$ are both affine on $J$. Then the map 
  \begin{equation*}
    f_0 \disjointunion f_1 \from J \times \{0\} \disjointunion J \times \{1\}
          \to K_n,
  \end{equation*}
  defined on a subset of the product simplicial complex $J \times [0,1]$,
  satisfies the conditions of Proposition~\todo{extending maps}, and so admits
  an affine extension $F \from \sd^{k'} (J \times [0,1]) \to K_n$. Let $h =
  i^n_\infty \composed F$, so $h$ is a homotopy from $i^n_\infty\composed f_0$
  to $i^n_\infty\composed f_1$ and it remains to prove the uniform continuity
  assertion.

@@

  Now, for $\eta \leq \eta_0$, there exists a maximal number $\kappa(\eta) \geq
  0$ such that for any $x \in \Delta^d$ and any $s$ and $t$ in $[0,1]$ with
  $\abs{s-t} \leq \eta$, the segment $\{x\} \times [s,t]$ is contained in a
  union of at most two $(d+1)$-simplices of
  $\sd^{\kappa(\eta)}\sd^{k'}(\Delta^d\times[0,1])$. Note that $\kappa(\eta)
  \to \infty$ as $\eta \to 0$. Now, if $N\in\naturals$ then a $(d+1)$-simplex of
  $\sd^{k^N}K_n$ is mapped by $i^n_{n+N}$ into a $(d+1)$-simplex of $K_{n+N}$,
  which is mapped by $i^{n+N}_\infty$ to a set of diameter at most
  $C_0a^{-n+N}$. Therefore, if $N$ is such that $\kappa(\eta) \geq k^N$ then
  for $x \in \boundary\Delta^{d+1}$ and $\abs{s-t} < \eta$, $\rho(h(x,s),
  h(x,t)) \leq 2C_0a^{-n-N}$. It follows that we may take $\mu(\eta)$ as follows:
  \begin{equation*}
    \mu(\eta) = 2C_0 a^{-\floor*{\frac{\log\kappa(\eta)}{\log k}}}.\qedhere
  \end{equation*}
\end{proof}

@@

\begin{proposition}\label{prop:uniform_homotopies}
  There exists a number $m_0$ and a function $\nu\from\posreals\to\posreals$ such
  that $\nu(\eta)\to0$ as $\eta\to0$ with the following property. Let
  $\gamma\from\boundary\Delta^{d+1}\to\boundary G$ be a continuous map. Let $n
  \geq m_0$ and let $f$ be an affine approximation to
  $p^\infty_n\composed\gamma$. Then there is a homotopy $h$ from $\gamma$ to
  $i^n_\infty\composed f$ such that for any $x \in \boundary\Delta^{d+1}$ and
  any $s$ and $t$ in $[0,1]$,
  \begin{equation*}
    \rho(h(x,s), h(x,t)) \leq \nu(\abs{s-t}) a^{-n}.
  \end{equation*}
\end{proposition}

@@

  Let $\gamma\from\boundary\Delta^{d+1}\to\boundary G$, $n\geq m_0$ and
  $f\from\boundary\Delta^{d+1}\to K_n$ be as in the proposition. Let $f_0 = f$,
  and for $m > 0$ let $f_m\from\boundary\Delta^{d+1}\to K_{n+m}$ be a
  simplicial approximation to $p^\infty_{n+m}\composed\gamma$. For $m \geq 0$
  let $\gamma_m = i^{m+n}_\infty\composed f_m$. For each $m\geq 0$ the
  following observations are immediate:
  \begin{enumerate}
    \item the map $f_m$ admits a factoring;
    \item the image of $f_m$ is contained in the $d$-skeleton of $K_{n+m}$, and
      so $i^{n+m}_{n+m+1}\composed f_m$ admits a factoring;
    \item by \cref{lem:name_this} we have $\newdist(i^{n+m}_{n+m+1}\composed
      f_m(x), f_{m+1}(x)) \leq C$ for all $x \in \boundary\Delta^{d+1}$.
  \end{enumerate}
  Therefore, for each $m$, \cref{prop:basic_homotopies} gives a homotopy $h_m$
  from $i^{n+m+1}_\infty\composed f_{m+1}$ to $i^{n+m}_\infty \composed f_m$
  satisfying a uniform continuity condition. Define an infinite concatenation
  of these homotopies:
  \begin{equation*}
    h(x,t) = 
    \begin{cases}
      \gamma(x) & \text{if } t = 0\\
      h_n\left(x,2^{m+1}\left(t-\frac{1}{2^{m+1}}\right)\right) 
              & \text{if $t \in \left[\frac{1}{2^{m+1}}, \frac{1}{2^m}\right]$
                    , $m\geq 0$}.
    \end{cases}
  \end{equation*}
  This function is clearly continuous except possibly on
  $\boundary\Delta^{d+1}\times\{0\}$. We aim to show it is also continuous
  there, but first we define the function $\nu$.

@@

  To begin, we bound the distance $\rho(h(x,0), h(x,t))$
  where $x \in \boundary \Delta^{d+1}$ and $t < 2^{-m}$. Let $C'$ be the
  constant appearing in \cref{lem:convergestoidentity}. Then 
  \begin{equation*}
    \rho(h(x,0), h(x,2^{-m})) \leq C'a^{-m-n}.
  \end{equation*}
  Furthermore, for any $m' \in \naturals$ and any $t \in [2^{-m'-1}, 2^{-m'}]$,
  \begin{equation*}
    \rho(h(x,t), h(x,2^{-m'})) \leq \mu(1)a^{-m'-n}.
  \end{equation*}
  It follows that for $t \leq 2^{-m}$,
  \begin{equation*}
    \rho(h(x,0), h(x,t)) \leq a^{-n}\left(C'a^{-m} + \sum_{m' \geq m}
          \mu(1)a^{-m'}\right).
  \end{equation*}

@@

  Therefore for $\eta \in (2^{-m-1}, 2^{-m}]$ we define
  \begin{equation*}
    \nu(\eta)  = \max\left\{
      C'a^{-m} + \sum_{m' \geq m} \mu(1)a^{-m'},
      \sum_{m' \geq \floor{m/2}} \mu(1) a^{-m'},
      \mu(2^{1-m/2})
    \right\},
  \end{equation*}
  and we see that $\nu(\eta) \to 0$ as $\eta \to 0$, and also that
  \begin{equation*}
    \rho(h(x,s), h(x,t)) \leq \nu(\abs{s-t})
  \end{equation*}
  for any $x \in \boundary\Delta^{d+1}$ and $s$ and $t$ in $[0,1]$.

  It remains to show that $h$ is continuous at $(x,0)$ for each $x$ in
  $\boundary\Delta^{d+1}$. Let $\epsilon > 0$ and take a neighbourhood $U$ of
  $x$ in $\boundary\Delta^{d+1}$ such that $\rho(\gamma(x), \gamma(y)) \leq
  \epsilon/2$ for $y$ in $U$. Let $\eta > 0$ be so that $\nu(\eta) <
  \epsilon/2$. Then for any $(y, t) \in U \times [0, \eta]$, $\rho(h(x,0),
  h(y,t)) < \epsilon$.
\end{proof}

@@

\begin{proof}
  To begin, let $\nu$ and $m_0 \geq n_0$ be as in
  \cref{prop:uniform_homotopies}. Let $r_0 =
  k_1a^{-m_0}$, so that, by \cref{lem:projectstoasimplex},
  $p^\infty_{m_0}\composed\gamma$ has image contained in the star of a single
  vertex of $K_{m_0}$ for any map $\gamma\from\boundary\Delta^{d+1}\to\boundary
  G$ with $\Diam\image(\gamma)\leq r_0$.

  Let $\gamma\from\boundary\Delta^{d+1}\to\boundary G$ be such a map. Let 
  \begin{equation*}
    m = \floor{\log_a(k_1/\Diam\image\gamma)} \geq m_0,
  \end{equation*}
  so that by \cref{lem:projectstoasimplex} $p^\infty_{m}\composed\gamma$ has
  image contained in a single vertex of $K_m$. Then a constant map
  $\boundary\Delta^{d+1} \to K_m$ is a simplicial approximation to
  $p^\infty_m\composed\gamma$; applying \cref{prop:uniform_homotopies} to
  this map immediately gives a homotopy $h$ from $\gamma$ to the constant map.

  Furthermore,
  \begin{align*}
    \Diam\image h &\leq \Diam\image\gamma + \sup_{x \in\boundary\Delta^{d+1}, s,t \in [0,1]} \rho(h(x,s), h(x,t)) \\
                  &\leq \Diam\image\gamma + a^{-m} \nu(1) \\
                  &\leq (1 + a\nu(1)/k_1) \Diam\image\gamma \qedhere
  \end{align*}
\end{proof}

>>>>

  We now bound the distance $\dist(y_n, y_{n+1})$. To start with, $\dist(y_n,
  i^n_{n+1} y_n) \leq 2\delta+1$. Let $R$ be $3\cdot 2^{k-1}$, so that any two
  vertices in the 1-skeleton of $\sd^k \Delta^2$ are joined by a path
  traversing at most $R$ edges. Let $\sigma$ be the minimal simplex of $K_n$
  containing $i^m_n(x)$. Then $y_n$ is a vertex of $\sigma$ and $y_{n+1}$ is a
  vertex of the image of $\sigma$ under $i^n_{n+1}$, so $\dist(i^n_{n+1} y_n,
  y_{n+1}) \leq RD$.  Putting this together,
  \begin{equation*}
      \dist(y_n, y_{n+1}) \leq 2\delta + 1 + RD
  \end{equation*}

@@

\begin{proposition}\label{prop:basic_homotopies}
  Let $C$ be a constant. Then there exists a number $m_0$ and a function $\mu
  \from \posreals\to\posreals$ such that $\mu(\eta)\to0$ as $\eta\to0$ with the following
  property. Let $n \geq m_0$ and let $f_0$ and $f_1$ be any maps from
  $\boundary\Delta^2$ to $K_n$ satisfying the following conditions:
  \begin{enumerate}
    \item the maps $f_0$ and $f_1$ are affine on some subdivision of
      $\boundary\Delta^2$;
    \item for any $x \in \boundary\Delta^2$, the distance 
      $\newdist(f_0(x), f_1(x))$ is at most $C$;
    \item each of $f_0$ and $f_1$ admits a factoring.
  \end{enumerate}
  Then there is a homotopy $h$ from $i^n_\infty\composed f_0$ to
  $i^n_\infty\composed f_1$ satisfying a uniform continuity condition: for any
  $x \in \boundary\Delta^2$ and any $s$ and $t$ in $[0,1]$,
  \begin{equation*}
    \rho(h(x,s), h(x,t)) \leq \mu(\abs{s-t})a^{-n}.
  \end{equation*}
\end{proposition}

@@

  Let $n \geq m_0$.  Let $J$ be a subdivision of $\boundary\Delta^2$ such
  that $f_0$ and $f_1$ are both affine on $J$. Then the map 
  \begin{equation*}
    f_0 \disjointunion f_1 \from J \times \{0\} \disjointunion J \times \{1\}
          \to K_n,
  \end{equation*}
  defined on a subset of the product simplicial complex $J \times [0,1]$,
  satisfies the conditions of Proposition~\todo{extending maps}, and so admits
  an affine extension $F \from \sd^{k'} (J \times [0,1]) \to K_n$. Let $h =
  i^n_\infty \composed F$, so $h$ is a homotopy from $i^n_\infty\composed f_0$
  to $i^n_\infty\composed f_1$ and it remains to prove the uniform continuity
  assertion.

@@

  Now, for $\eta \leq \eta_0$, there exists a maximal number $\kappa(\eta) \geq
  0$ such that for any $x \in \Delta^d$ and any $s$ and $t$ in $[0,1]$ with
  $\abs{s-t} \leq \eta$, the segment $\{x\} \times [s,t]$ is contained in a
  union of at most two $(d+1)$-simplices of
  $\sd^{\kappa(\eta)}\sd^{k'}(\Delta^d\times[0,1])$. Note that $\kappa(\eta)
  \to \infty$ as $\eta \to 0$. Now, if $N\in\naturals$ then a $(d+1)$-simplex of
  $\sd^{k^N}K_n$ is mapped by $i^n_{n+N}$ into a $(d+1)$-simplex of $K_{n+N}$,
  which is mapped by $i^{n+N}_\infty$ to a set of diameter at most
  $C_0a^{-n+N}$. Therefore, if $N$ is such that $\kappa(\eta) \geq k^N$ then
  for $x \in \boundary\Delta^2$ and $\abs{s-t} < \eta$, $\rho(h(x,s),
  h(x,t)) \leq 2C_0a^{-n-N}$. It follows that we may take $\mu(\eta)$ as follows:
  \begin{equation*}
    \mu(\eta) = 2C_0 a^{-\floor*{\frac{\log\kappa(\eta)}{\log k}}}.\qedhere
  \end{equation*}
\end{proof}

@@

\begin{proposition}\label{prop:uniform_homotopies}
  There exists a number $m_0$ and a function $\nu\from\posreals\to\posreals$ such
  that $\nu(\eta)\to0$ as $\eta\to0$ with the following property. Let
  $\gamma\from\boundary\Delta^2\to\boundary G$ be a continuous map. Let $n
  \geq m_0$ and let $f$ be an affine approximation to
  $p^\infty_n\composed\gamma$. Then there is a homotopy $h$ from $\gamma$ to
  $i^n_\infty\composed f$ such that for any $x \in \boundary\Delta^2$ and
  any $s$ and $t$ in $[0,1]$,
  \begin{equation*}
    \rho(h(x,s), h(x,t)) \leq \nu(\abs{s-t}) a^{-n}.
  \end{equation*}
\end{proposition}

@@

  Let $\gamma\from\boundary\Delta^2\to\boundary G$, $n\geq m_0$ and
  $f\from\boundary\Delta^2\to K_n$ be as in the proposition. Let $f_0 = f$,
  and for $m > 0$ let $f_m\from\boundary\Delta^2\to K_{n+m}$ be a
  simplicial approximation to $p^\infty_{n+m}\composed\gamma$. For $m \geq 0$
  let $\gamma_m = i^{m+n}_\infty\composed f_m$. For each $m\geq 0$ the
  following observations are immediate:
  \begin{enumerate}
    \item the map $f_m$ admits a factoring;
    \item the image of $f_m$ is contained in the $d$-skeleton of $K_{n+m}$, and
      so $i^{n+m}_{n+m+1}\composed f_m$ admits a factoring;
    \item by \cref{lem:name_this} we have $\newdist(i^{n+m}_{n+m+1}\composed
      f_m(x), f_{m+1}(x)) \leq C$ for all $x \in \boundary\Delta^2$.
  \end{enumerate}
  Therefore, for each $m$, \cref{prop:basic_homotopies} gives a homotopy $h_m$
  from $i^{n+m+1}_\infty\composed f_{m+1}$ to $i^{n+m}_\infty \composed f_m$
  satisfying a uniform continuity condition. Define an infinite concatenation
  of these homotopies:
  \begin{equation*}
    h(x,t) = 
    \begin{cases}
      \gamma(x) & \text{if } t = 0\\
      h_n\left(x,2^{m+1}\left(t-\frac{1}{2^{m+1}}\right)\right) 
              & \text{if $t \in \left[\frac{1}{2^{m+1}}, \frac{1}{2^m}\right]$
                    , $m\geq 0$}.
    \end{cases}
  \end{equation*}
  This function is clearly continuous except possibly on
  $\boundary\Delta^2\times\{0\}$. We aim to show it is also continuous
  there, but first we define the function $\nu$.

@@

  To begin, we bound the distance $\rho(h(x,0), h(x,t))$
  where $x \in \boundary \Delta^2$ and $t < 2^{-m}$. Let $C'$ be the
  constant appearing in \cref{lem:convergestoidentity}. Then 
  \begin{equation*}
    \rho(h(x,0), h(x,2^{-m})) \leq C'a^{-m-n}.
  \end{equation*}
  Furthermore, for any $m' \in \naturals$ and any $t \in [2^{-m'-1}, 2^{-m'}]$,
  \begin{equation*}
    \rho(h(x,t), h(x,2^{-m'})) \leq \mu(1)a^{-m'-n}.
  \end{equation*}
  It follows that for $t \leq 2^{-m}$,
  \begin{equation*}
    \rho(h(x,0), h(x,t)) \leq a^{-n}\left(C'a^{-m} + \sum_{m' \geq m}
          \mu(1)a^{-m'}\right).
  \end{equation*}

@@

  Therefore for $\eta \in (2^{-m-1}, 2^{-m}]$ we define
  \begin{equation*}
    \nu(\eta)  = \max\left\{
      C'a^{-m} + \sum_{m' \geq m} \mu(1)a^{-m'},
      \sum_{m' \geq \floor{m/2}} \mu(1) a^{-m'},
      \mu(2^{1-m/2})
    \right\},
  \end{equation*}
  and we see that $\nu(\eta) \to 0$ as $\eta \to 0$, and also that
  \begin{equation*}
    \rho(h(x,s), h(x,t)) \leq \nu(\abs{s-t})
  \end{equation*}
  for any $x \in \boundary\Delta^2$ and $s$ and $t$ in $[0,1]$.

  It remains to show that $h$ is continuous at $(x,0)$ for each $x$ in
  $\boundary\Delta^2$. Let $\epsilon > 0$ and take a neighbourhood $U$ of
  $x$ in $\boundary\Delta^2$ such that $\rho(\gamma(x), \gamma(y)) \leq
  \epsilon/2$ for $y$ in $U$. Let $\eta > 0$ be so that $\nu(\eta) <
  \epsilon/2$. Then for any $(y, t) \in U \times [0, \eta]$, $\rho(h(x,0),
  h(y,t)) < \epsilon$.
\end{proof}

@@

\begin{proof}
  To begin, let $\nu$ and $m_0 \geq n_0$ be as in
  \cref{prop:uniform_homotopies}. Let $r_0 =
  k_1a^{-m_0}$, so that, by \cref{lem:projectstoasimplex},
  $p^\infty_{m_0}\composed\gamma$ has image contained in the star of a single
  vertex of $K_{m_0}$ for any map $\gamma\from\boundary\Delta^2\to\boundary
  G$ with $\Diam\image(\gamma)\leq r_0$.

  Let $\gamma\from\boundary\Delta^2\to\boundary G$ be such a map. Let 
  \begin{equation*}
    m = \floor{\log_a(k_1/\Diam\image\gamma)} \geq m_0,
  \end{equation*}
  so that by \cref{lem:projectstoasimplex} $p^\infty_{m}\composed\gamma$ has
  image contained in a single vertex of $K_m$. Then a constant map
  $\boundary\Delta^2 \to K_m$ is a simplicial approximation to
  $p^\infty_m\composed\gamma$; applying \cref{prop:uniform_homotopies} to
  this map immediately gives a homotopy $h$ from $\gamma$ to the constant map.

  Furthermore,
  \begin{align*}
    \Diam\image h &\leq \Diam\image\gamma + \sup_{x \in\boundary\Delta^2, s,t \in [0,1]} \rho(h(x,s), h(x,t)) \\
                  &\leq \Diam\image\gamma + a^{-m} \nu(1) \\
                  &\leq (1 + a\nu(1)/k_1) \Diam\image\gamma \qedhere
  \end{align*}
\end{proof}
